---
output: github_document
always_allow_html: yes
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# getRemoteData

<!-- badges: start -->
<!-- badges: end -->

`getRemoteData` is an R package that offers a common framework to query and import remote data (i.e. data stored on the cloud) from heterogeneous sources. Overall, this package attempts to **facilitate** and **speed-up** the painfull and time-consuming **data import / download** process for some well-known and widely used environmental / climatic data (e.g. [MODIS](https://modis.gsfc.nasa.gov/), [VNP](https://lpdaac.usgs.gov/search/?query=VNP&page=2), [GPM](https://www.nasa.gov/mission_pages/GPM/main/index.html), etc.) as well as other sources (e.g. [VIIRS DNB](https://ngdc.noaa.gov/eog/viirs/download_dnb_composites.html), etc.). You will take the best of `getRemoteData` if you work at **local to regional** spatial scales, i.e. typically from few decimals to a decade squared degrees. For larger areas, other packages might be more relevant (see section [Other relevant packages](#other-relevant-packages) ).

**Why such a package ?**  

Modeling an ecological phenomenon (e.g. species distribution) using environmental data (e.g. temperature, rainfall) is quite a common task in ecology. The data analysis workflow generally consists in : 

- importing, tidying and summarizing various environmental data at geographical locations and dates of interest ;
- creating explicative / predictive models of the phenomenon using the environmental data.

Data of interest for a specific study are usually heterogeneous (various sources, formats, etc.). Downloading long time series of several environmental data "manually" (e.g. through user-friendly web portals) is time consuming, not reproducible and prone to errors. In addition, when downloaded manually, spatial datasets might cover quite large areas, or include many dimensions (e.g. the multiple bands for a MODIS product). If your aera of interest is smaller or if you do not need all the dimensions, why downloading the whole dataset ? Whenever possible (i.e. made possible by the data provider - check section [Behind the scene... how it  works](#behind-the-scene-...-how-it-works)), `getRemoteData` enables to download the data strictly for your region and dimensions of interest.

Finally, getRemoteData relies as much as possible on open and standard data access protocols (eg. [OPeNDAP](https://en.wikipedia.org/wiki/OPeNDAP)), which makes it (and by extension, your script) less vulnerable to external changes than packages or applications relying on APIs.

**When should you use `getRemoteData` ?**

`getRemoteData` can hopefully help if you recognize yourself in one or more of the following points :

- work at a local to regional spatial scale ; 
- need to import data from various sources (e.g. MODIS, GPM, etc.) ;
- are interested in importing long climatic / environmental time-series ; 
- have a slow internet connection ; 
- care about the digital environmental impact of your work.

`getRemoteData` is developed in the frame of PhD project, and the sources of data implemented in the package are hence those that I use in my work. Sources of data are mostly environmental / climatic data, but not exclusively. Have a look at the function `getAvailableDataSources` to check which sources are already implemented ! 

## Installation

You can install the development version of `getRemoteData` from [GitHub](https://github.com/) with: 

``` r
# install.packages("devtools")
devtools::install_github("ptaconet/getRemoteData")
```

## Get the data sources implemented in `getRemoteData`

You can get the data sources/collections downloadable with `getRemoteData` and details about each of them with : 

```{r getAvailableDataSources_call, eval = FALSE, echo=T, warning=F, message=F}
getRemoteData::getAvailableDataSources()
```
```{r, eval = TRUE, echo=F}
library(kableExtra)
getRemoteData::getAvailableDataSources() %>%
  dplyr::select(-citation) %>%
  kable() %>%
  kable_styling()
```

## Example

Say you want to download  over a 3500km^2^ region of interest:

- a 40 days time series of [MODIS Terrra Land Surface Temperature (LST)](https://dx.doi.org/10.5067/MODIS/MOD11A1.006) (daily time resolution);
- the same 40 days times series of [Global Precipitation Measurement (GPM)](https://doi.org/10.5067/GPM/IMERGDF/DAY/06) (daily time resolution) :

```{r example, eval = FALSE}
library(getRemoteData)
library(sf)
library(purrr)
# Read the region of interest as a sf object. Here : the Korhogo area in Côte D'Ivoire
roi<-sf::st_read(system.file("extdata/ROI_example.kml", package = "getData"),quiet=T)
# Set-up your time frame of interest. 
time_frame<-c("2017-05-01","2017-06-10")
# Set-up your credentials to EarthData
username_EarthData<-"my.earthdata.username"
password_EarthData<-"my.earthdata.username"
# Download the MODIS LST TERRA daily products in the current working directory
# Setting the argument 'download' to FALSE will return the URLs of the products, without downloading them 
dl_modis<-getRemoteData::getData_modis(timeRange = time_frame,
                                     roi = roi,
                                     collection="MOD11A1.006",
                                     dimensions=c("LST_Day_1km","LST_Night_1km"),
                                     download = T,
                                     destFolder=getwd(),
                                     username=username_EarthData,
                                     password=password_EarthData,
                                     parallelDL=T #setting to F will download the data linearly
                                     )
head(dl_modis)
# Download the GPM daily products in the current working directory
dl_gpm<-getRemoteData::getData_gpm(timeRange = time_frame,
                                     roi = roi,
                                     collection="GPM_3IMERGDF.06",
                                     dimensions=c("precipitationCal"),
                                     download = T,
                                     destFolder=getwd(),
                                     username=username_EarthData,
                                     password=password_EarthData,
                                     parallelDL=T
                                     )
head(dl_gpm)

# Get the data downloaded as a list of rasters
rasts_modis<-dl_modis$destfile %>%
  purrr::map(~getRemoteData::prepareData_modis(.,"LST_Day_1km")) %>%
  set_names(dl_modis$name)

rasts_gpm<-dl_gpm$destfile %>%
  purrr::map(~getRemoteData::prepareData_gpm(.,"precipitationCal")) %>%
  set_names(dl_gpm$name)
```

The functions of `getRemoteData` all work the same way : 

- `timeRange` is your date / time frame of interest (eventually including hours for the data with less that daily resolution) ;
- `roi` is your area of interest (as an `sf` object, either point or polygon) ;
- `destfolder` is the data destination folder ;
- by default, the function does not download the dataset. It returns a data.frame with the URL(s) to download the dataset(s) of interest given the input arguments. To download the data, set the *download* argument to TRUE ;
- other arguments are specific to each data product (e.g. `collection`, `dimensions`,`username`,`password`)

Absence of the `timeRange` (resp. `roi`) arguments in a function means that the data of interest do not have any temporal (resp. spatial) dimension.

Have a look at the vignette [Automatic extraction of spatial-temporal environmental data within buffers around sampling points](https://ptaconet.github.io/malamodpkg/articles/import_tidy_transform_envdata.html) to get an example of what you can do with `getRemoteData` !

## Current limitations

The package is at a very early stage of development. Here are some of the current limitations and ideas of future developments : 

- MODIS data cannot be downloaded if your area of interest covers multiple MODIS tiles (for an overview of MODIS tiles go [here](https://modis.ornl.gov/files/modis_sin.kmz));

## Behind the scene... how it works

As much as possible, when implemented by the data providers, `getRemoteData` uses open and standard data access protocols to download the data. These standard protocols enable to filter the data directly at the downloading phase. Filters can be spatial, temporal or dimensional. Example of widely-used standard data access protocols for geospatial timeseries are [OGC WFS](https://en.wikipedia.org/wiki/Web_Feature_Service) or [OPeNDAP](https://en.wikipedia.org/wiki/OPeNDAP). 
If long time series are queried, `getRemoteData` enables to speed-up the downloading time by parallelizing it.

## Other relevant packages

- [`getSpatialData`](http://jxsw.de/getSpatialData/)
- [`MODIS`] and [`MODISTools`] and [`MODISTsp`]
- GPM ?

