---
title: "Efficient extraction of environmental/climatic spatial-temporal series over small-scale areas"
author: "Paul Taconet"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Efficient extraction of environmental/climatic spatial-temporal series over small-scale areas}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Objectives

In this vignette we show how to efficiently extract time-space statistics of various climatic / environmental variables at given dates and locations with the R packages `getData`, `purrr` and `furrr`.

The overall objective is to extract : within a given **buffer size** (e.g. 1000 m) around a given **point location** (e.g. lat -10.5, lon -5.2) at a given **date** (e.g. 2017-05-01), a **summary** (e.g. mean) of a given **variable** (e.g. land surface temperature). This workflow enables to set as input parameters : 

- Various {dates, locations} couples ;
- Various buffer sizes ;
- Various climatic / environmental data sources (e.g NASA MODIS products, GPM, ERA-5, etc.) ;
- A laglength (i.e. number of days before the dates of interest for which also to extract the variable)

Example : 

Given the following set of {dates, locations} :


```{r setup}
library(getData)
```
