---
title: "1. Get started "
author: "Paul Taconet"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true") # vignette will not be executed when tested on the cran
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  purl = NOT_CRAN
  )
```

The example below is a full time-series data download, import and plot workflow in 4 steps : get the URL with the package `getremotedata`, download with the package `httr`, import and plot with the package `raster`. Packages `purrr` and `sf` are also used.

Data of interest are : 

* Elevation ([SRTM DEM](https://dx.doi.org/10.5067/MEASURES/SRTM/SRTMGL1.003)) ;
* Daily rainfall ([TAMSAT](http://doi.org/10.1038/sdata.2017.63))
* Monthly nighttime lights ([VIIRS_DNB_MONTH](https://doi.org/10.5067/VIIRS/VNP46A1.001))
* Daily moonlight ([MIRIADE](http://vo.imcce.fr/webservices/miriade/))
* Hourly wind speed and direction ([ERA5](https://doi.org/10.24381/cds.adbb2d47))

Region of interest : 

```{r, echo=F, fig.height=5, fig.width=3, eval=T }
require(mapview)
roi <- sf::st_as_sf(data.frame(geom="POLYGON ((-5.82 9.54, -5.42 9.55, -5.41 8.84, -5.81 8.84, -5.82 9.54))"),wkt="geom",crs = 4326)
mapview::mapview(roi,legend=F)
```

**Preliminary step : call the packages and set-up the region and the time range of interest**

```{r setup, message=F}
require(getremotedata)
require(sf)
require(raster)
require(magrittr)
require(purrr)
require(httr)
```

```{r example_setup, eval=T ,message=F, warning=F}
roi <- st_as_sf(data.frame(geom = "POLYGON ((-5.82 9.54, -5.42 9.55, -5.41 8.84, -5.81 8.84, -5.82 9.54))"), wkt = "geom", crs = 4326)
time_range <- as.Date(c("2017-01-01","2017-01-30"))
```

**1. Get the URLs of the data for a given ROI and time frame** 

```{r example_geturl, eval=T ,message=F, warning=F}
roi <- st_as_sf(data.frame(geom = "POLYGON ((-5.82 9.54, -5.42 9.55, -5.41 8.84, -5.81 8.84, -5.82 9.54))"), wkt = "geom", crs = 4326)
time_range <- as.Date(c("2017-01-01","2017-01-30"))

# SRTM
strm_urls <- grd_get_url(collection = "SRTMGL1.003", roi = roi)
print(str(strm_urls))

# TAMSAT
tamsat_urls <- grd_get_url(collection = "TAMSAT",variables = c("daily_rainfall_estimate"), time_range = time_range)
print(str(tamsat_urls))

# VIIRS_DNB_MONTH
viirsdnb_urls <- grd_get_url(collection = "VIIRS_DNB_MONTH",variables = c("Monthly_AvgRadiance"), roi = roi, time_range = time_range)
print(str(viirsdnb_urls))

# MIRIADE
imcce_urls <- grd_get_url(collection = "MIRIADE", roi = roi, time_range = time_range)
print(str(imcce_urls))

# ERA5
# ERA5 is an hourly database, so we keep only 1 day (i.e. not the whole month)
era5_urls <- grd_get_url(collection = "ERA5", variables = c("10m_u_component_of_wind","10m_v_component_of_wind"), roi = roi, time_range = as.Date(c("2017-01-01","2017-01-02")))
```

**2. Download the data**

```{r example_download, eval=F ,message=F, warning=F}
# Create directories if they do not exist
unique(dirname(data_to_dl$destfile)) %>% lapply(dir.create,recursive = TRUE)

# SRTM
# Login to Earthdata servers is needed to download STRM data. To create an account go to : https://urs.earthdata.nasa.gov/.
# Here we have stored our credentials in local environment variables
username <- Sys.getenv("earthdata_un")
password <- Sys.getenv("earthdata_pw")
srtm_dl <- map2(strm_urls$url,strm_urls$destfile,~GET(url = .x, write_disk(.y), progress() , authenticate(username,password)))

# TAMSAT
tamsat_dl <- map2(tamsat_urls$url,tamsat_urls$destfile,~GET(url = .x, write_disk(.y), progress()))

# VIIRS_DNB_MONTH
viirsdnb_dl <- map2(viirsdnb_urls$url,viirsdnb_urls$destfile,~GET(url = .x, write_disk(.y), progress()))

# MIRIADE
imcce_dl <- map2(imcce_urls$url,imcce_urls$destfile,~GET(url = .x, write_disk(.y), progress()))

# ERA5
# For ERA5 we must use a specific function to download the data
era5_dl <- grd_download_data_era5(era5_urls)
```

**3. Import the data in R**

```{r example_import, eval=T ,message=F, warning=F}
# SRTM 
rast_srtm <- strm_urls$destfile %>%
  map(~unzip(., exdir = dirname(.))) %>%
  map(~raster(.)) %>%
  do.call(merge,.) %>%
  crop(roi)

# TAMSAT
rast_tamsat <- tamsat_urls$destfile %>%
  map(~raster(.)) %>%
  map(~crop(.,roi)) %>%
  brick()
names(rast_tamsat) <- tamsat_urls$time_start

# VIIRS_DNB_MONTH
rasts_viirs <- viirsdnb_urls$destfile %>%
  map(~raster(.)) %>%
  brick(.)
names(rasts_viirs) <- viirsdnb_urls$time_start

# MIRIADE
miriade_dfs <- imcce_urls$destfile %>%
  map(~read.csv(.,skip=10))

# ERA5
rasts_era5 <- era5_urls$destfile %>%
  map(~raster(., varname = "u10")) %>%
  brick(.)
names(rasts_era5) <- era5_urls$time_start


rast_srtm
rast_tamsat
rasts_viirs
print(str(miriade_dfs[1:3]))
rasts_era5
```

**4. Display / Plot**

```{r plot , fig.fullwidth = TRUE, fig.height = 10, fig.width = 9,fig.align = 'center', eval=T}
# SRTM 
plot(rast_srtm)

# TAMSAT
plot(rast_tamsat)

# VIIRS_DNB_MONTH
plot(rasts_viirs)

# MIRIADE
print(str(miriade_dfs[1:3]))

# ERA5
plot(rasts_era5)
```

